Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-03 examples.tex
+

\newcommand{\fpaddr}{\texttt{fpaddr}}
\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
\newcommand{\sumapacesfull}[2]{
  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
}

@@

Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).

The crucial step is actually updating the L1 entry,
via the virtual address
(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
allocated physical page address (\textsf{fpaddr}) in Line 18.
After this write, we have (Specification Line 21)
\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]

At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
physical pointsto
($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
Since we know $\texttt{addr\_L1 }(\vaddr,
\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, we can
connect L4-L2 points-to assertions with 
 $\texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})$ to obtain the
complete physical table-walk assertion (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
fpaddr)) in Line 25.

@@

Within \textsf{get\_next\_table} first checks if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$). To be able to do so, it needs to load the entry value (a 64-bits word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.

@@

$\specline{ \ulcorner \Xi !! \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$

@@

@@

As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 7 in Figure \ref{fig:calltopteinitialize}). We need to construct this virtual mapping for \textsf{entry}.
\subsubsection{Identity Mappings}
\label{subsec:identitymappings}
Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\begin{figure*}
\footnotesize
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \\
                           \qquad\qquad\qquad\qquad\qquad\qquad          \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}

>>>>

\newcommand{\fpaddr}{\texttt{fpaddr}}
\newcommand{\specline}[1]{{\color{blue}\left\{#1\right\}}}
\newcommand{\sumapacesfull}[2]{
  \ownGhost\gammaPreds{\authfull{\singletonMap{#1}{#2}}}
}
\todo{ismail change the following spec base on the changes you have in the code}

@@

%In the precondition, we see Line 12 allocates a fresh page-aligned, zero-initialized page  (at \textsf{fpaddr}),
%returning a pre-filled PTE entry in \textsf{rax} ($+3$ sets the lower 2 bits).

The crucial step is actually updating the L1 entry,
via the virtual address
(\textsf{pte\_addr}) known to translate to the appropriate physical address, in our example the L1
table entry address ($\textsf{addr\_L1}(\textsf{va, l1e})$), to hold the freshly
allocated physical page address (\textsf{fpaddr}) in Line X.
After this write, we have (Specification Line Y)
\[\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr}  \qquad\qquad (\textsf{by the PTE variant of Rule}~\TirNameStyle{WriteToVirtMemFromReg})\]

At this point in the proof,  $\texttt{pte\_addr} \mapsto_{\texttt{vpte}}  \; \paddr\; \textsf{fpaddr})$ 
contains  full ownerhsip of the L1 table entry physical pointsto $\paddr
\mapsto_{\mathsf{p}} \textsf{fpaddr}$ . Then, we can split the full-ownership of that
physical pointsto
($\textsf{fpaddr}\mapsto_{\textsf{p}} \;\textsf{wzero 64}$) to obtain
fractional ownership on it appropriate for a single address's share of an L1 entry (q4).
Since we know $\texttt{addr\_L1 }(\vaddr,
\entryo) \mapsto_{\mathsf{p}} \{q4\} \;\textsf{fpaddr}$) from
$\mapsto_{\textsf{vpte}}$, after the split of the PTE points-to, and, thanks to the post-condition of the page-table walk (Specification of \textsf{pte\_walkpgdir} in Figure \ref{}), we know the existence of L4-L2 mappings reaching to the L1 exist, so that we can add the last mapping $\texttt{fpaddr} \mapsto_{p} (\texttt{wzero 64})$ to obtain the complete table-walk.
\todo{with the virtual-pte-pointsto assertions (L4\_L1\_PointsTo($\vaddr$ l4e l3e l2e
fpaddr)) in Line 25}

@@

Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry}$) is present or not. To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}) and check that \textit{presence} bit of the entry is set or not.

@@

$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$

@@

$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi\setminus \{ \mathsf{entry} \}\cup\{\mathsf{nextpaddr}\} \;!!\; \mathsf{entry} = \mathsf{Some Z} \urcorner \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$

@@

As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 13 in Figure \ref{fig:calltopteinitialize}). We need to construct this virtual mapping for the physical address \textsf{entry}.
\subsubsection{Identity Mappings}
\label{subsec:identitymappings}
Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\begin{figure*}
\footnotesize
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \bigast{(\paddr) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
